---
title: "Regression Models For Motor Trends"
author: "Kumar Shaket"
date: "01/03/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Executive Summary
In this report, we will examine the mtcars data set and explore how miles per gallon (MPG) is affected by different variables. In particularly, we will answer the following two questions: 

(1) *Is an automatic or manual transmission better for MPG* 
(2) *Quantify the MPG difference between automatic and manual transmissions*

# Exploratory Analysis
## Loading required libraries and dataset=mtcars
```{r}
library(ggplot2)
library(datasets)
data(mtcars)
head(mtcars,3)
```
## Transforming certain variables to factor variables
```{r}
mtcars$cyl <- factor(mtcars$cyl)
mtcars$vs <- factor(mtcars$vs)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
mtcars$am <- factor(mtcars$am, labels = c("Automatic","Manual"))
head(mtcars,5)
```
# Regression Analysis
We see that manual transmission is better for mpg than automatic transmission for cars as seen below.
```{r}
aggregate(mpg ~ am, data=mtcars,mean)
```
We hypothesize that automated transmission has approx 7 mpg lower than manual transmission. To determine if there is significant difference, let's do t.test.
```{r}
t_automatic <- mtcars[mtcars$am=="Automatic",]
t_manual <- mtcars[mtcars$am=="Manual",]
t.test(t_automatic$mpg,t_manual$mpg)
```
The p-value is 0.001374, thus we can state this is a significant difference. Now to quantify this, we will build 

## Linear Models

### Model1
```{r}
mt1 <- lm(mpg ~ am,mtcars)
summary(mt1)
```
This shows us that the average MPG for automatic is 17.1 MPG, while manual is 7.2 MPG higher via boxplot(Appendix - Plot 1) 

The R2 value is 0.36 thus telling us  model only explains us 36% of the variance. As a result, we need to build a multivariate linear regression.

### Model2

The new model will use the other variables to make it more accurate. We explore the other variable via a pairs plot (Appendix - Plot 2) to see how all the variables correlate with mpg. 
```{r}
mt2 <- lm(mpg ~ am + cyl + disp + hp + wt,mtcars)
summary(mt2)
```
From this we see that cyl, disp, hp, wt have the strongest correlation with mpg. 

We will compare both the models using anova function.
```{r}
anova(mt1,mt2)
```
This results in a p-value of 8.637e-08, and we can claim the mt2 model is significantly better than our mt1 linear model.

```{r}
summary(mt2)
```
The model explains 86.64% of the variance and as a result, cyl, disp, hp, wt did affect the correlation between mpg and am. 

#### Thus, we can say the difference between automatic and manual transmissions is 1.81 MPG.

# Appendix

## Plot 1 : Boxplot of MPG by transmission type
```{r}
boxplot(mpg ~ am,data = mtcars, col = (c("red","blue")), ylab = "Miles Per Gallon", xlab = "Transmission Type")
```

## Plot 2 - Pairs plot for the data set
```{r}
pairs(mpg ~ ., data = mtcars)
```

We double-check the residuals for non-normality (Appendix - Plot 3) and can see they are all normally distributed.

## Plot 3 - Check residuals
```{r}
par(mfrow = c(2,2))
plot(mt2)
```